---
title: "class19.Rmd"
author: "Anu Chaparala"
date: "11/30/2021"
output: pdf_document
---

# Section 1. Proportion of G/G in Population

Downloaded csv file from https://uswest.ensembl.org/Homo_sapiens/Variation/Sample?db=core;r=17:39825096-39965097;v=rs8067378;vdb=variation;vf=105535077#373531_tablePanel

Read CSV file

```{r}
mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head(mxl)
```

```{r}
table(mxl$Genotype..forward.strand.)
```

```{r}
table(mxl$Genotype..forward.strand.) / nrow(mxl)
```

Section 4: Population Scale Analysis
[HOMEWORK]
One sample is obviously not enough to know what is happening in a population. You are interested in assessing genetic differences on a population scale. So, you processed about ~230 samples and did the normalization on a genome level. Now, you want to find whether there is any association of the 4 asthma-associated SNPs (rs8067378...) on ORMDL3
expression.

```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
head(expr)
```
>Q13: Read this file into R and determine the sample size for each genotype and their corresponding median expression levels for each of these genotypes.

Median Exp Levels
A/A: 32
A/G: 25
G/G: 20

Sample Size
A/A: 108
A/G: 233
G/G: 121

```{r}
nrow(expr)
table(expr$geno)

#summary(plot$data)

```


```{r}
library(ggplot2)
```


Make box plot

```{r}
give.n <- function(x){
  return(c(y = median(x)*1.05, label = length(x))) 
  # experiment with the multiplier to find the perfect position
}

plot <- ggplot(expr, aes(geno, exp, fill=geno)) + geom_boxplot(notch=TRUE)  + stat_summary(fun.data = give.n, geom = "text", fun.y = median)

plot
```






