---
title: "class17.Rmd"
author: "Anu Chaparala"
date: "11/23/2021"
output:
  pdf_document:
    latex_engine: xelatex
---

##Background

As we approach a period of travel and larger gatherings, lets have a look at vaccination rates across the US. 
We will take data from the data.ca.gov site under “Statewide COVID-19 Vaccines Administered by ZIP Code” CSV file from: https://data.ca.gov/dataset/covid-19-vaccine-progress-dashboard-data-by-zip-code. 

##Dowload Data
```{r}
# Import vaccination data
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
head(vax)

```

## Ensure the date column is useful

>Q1. What column details the total number of people fully vaccinated? 

persons_fully_vaccinated

>Q2. What column details the Zip code tabulation area? 

zip_code_tabulation_area

>Q3. What is the earliest date in this dataset? 

2021-01-05

```{r}
#look at first and last dates in column
#vax$as_of_date
```


>Q4. What is the latest date in this dataset?

2021-11-16

```{r}
#install.packages("lubridate")
library(lubridate)
```

```{r}
today()
```


```{r}
#Specify that we are using the Year-month-day format
vax$as_of_date <- ymd(vax$as_of_date)
```

```{r}
#reverse_zipcode(c('92037', "92109"))
```

Now I can do useful math with dates more easily: 

```{r}
today() - vax$as_of_date[1]
```

How many days since the first entry?
```{r}
today() - vax$as_of_date[nrow(vax)]
```

>Q5. How many numeric columns are in this dataset? 

9

>Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column? 

8256

>Q7. What percent of persons_fully_vaccinated values are missing? 

10.2% or, 10%

>Q8. [Optional]: Why might this data be missing?

under-reporting, non-compliance during data collection, user error

```{r}
sum( is.na(vax$persons_fully_vaccinated) )

#sum( is.na(vax$persons_fully_vaccinated) )/ sum(vax$persons_fully_vaccinated, na.rm=TRUE)

skimr::skim(vax)

1-0.898

```

How many days does this set span?
```{r}
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```

>Q9. How many days have passed since the last update of the dataset? 

```{r}
today()-vax$as_of_date[nrow(vax)]
```
7

>Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)?

```{r}
length(unique(vax$as_of_date))
```
46

This makes sense.
```{r}
46*7
```

## Working with Zip Codes

```{r}
#install.packages("zipcodeR")
#library(zipcodeR)
```

Focus on  San Diego County
```{r}
table(vax$county)
```

We will subset with base R

```{r}
inds <- vax$county == "San Diego"

head(vax[inds,])

```

But let's use the **dyplr** package and it's **filer()** function: 
```{r}
library(dplyr)

sd <- filter(vax, county == "San Diego")

head(sd)

```
How many entries are there for San Diego county?
```{r}
nrow(sd)
```
Using dplyr is often more convenient when we are subsetting across multiple criteria - for example all San Diego county areas with a population of over 10,000.

```{r}
sd.10 <- filter(vax, county == "San Diego" &
                age5_plus_population > 10000)
```


>Q11. How many distinct zip codes are listed for San Diego County? 

```{r}
length(unique(sd$zip_code_tabulation_area))
```

107

>Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset?

```{r}
ind <- which.max(sd$age12_plus_population)
sd[ind,]
```
92154


>What is the population in the 92037 ZIP code area?

```{r}
filter(sd, zip_code_tabulation_area == "92037")[1,]
```



>Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2021-11-09”? 

67.28% average

```{r}
sd.now <- filter(sd, as_of_date=="2021-11-09")

head(sd.now)

mean(sd.now$percent_of_population_fully_vaccinated, na.rm=TRUE)
```


We can look at the 6-number summary

```{r}
summary(sd.now$percent_of_population_fully_vaccinated)
```

>Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2021-11-09”?

Using base R graphics
```{r}
hist(sd.now$percent_of_population_fully_vaccinated)
```


Using ggplot
```{r}
library(ggplot2)

ggplot(sd.now) + aes(percent_of_population_fully_vaccinated) + geom_histogram(bins=15)
```


What about 92037 - UCSD/La Jolla?

```{r}
filter(sd.now, zip_code_tabulation_area == "92037")
```

91% fully vaccinated!

```{r}
ucsd <- filter(sd, zip_code_tabulation_area=="92037")
ucsd[1,]$age5_plus_population
```


>Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area:

```{r}
ggplot(ucsd) +
 aes(as_of_date, percent_of_population_fully_vaccinated) + 
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated")
```

Let’s return to the full dataset and look across every zip code area with a population at least as large as that of 92037 on as_of_date “2021-11-16”.

```{r}
# Subset to all CA areas with a population as large as 92037
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2021-11-16")

head(vax.36)
```


>Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2021-11-16”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?

```{r}
mean1 <- mean(vax.36$percent_of_population_fully_vaccinated)
mean1
```

```{r}
ggplot(ucsd) +
 aes(as_of_date, percent_of_population_fully_vaccinated) + 
  geom_point() +
  geom_line(group=1) +
  geom_hline(yintercept=0.6629812, color = "red", linetype = "dashed") +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated")
```

>Q17. What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2021-11-16”?

```{r}
summary(vax.36$percent_of_population_fully_vaccinated)
```

>Q18. Using ggplot generate a histogram of this data.

```{r}
ggplot(vax.36) + 
  aes(percent_of_population_fully_vaccinated) + 
  geom_histogram(bins=25) + 
  labs(x="Percent Vaccinated", y="Count")
```

>Q19. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?

92109-above 92040-below

```{r}
vax %>% filter(as_of_date == "2021-11-16") %>%  
  filter(zip_code_tabulation_area=="92109") %>%
  select(percent_of_population_fully_vaccinated)

vax %>% filter(as_of_date == "2021-11-16") %>%  
  filter(zip_code_tabulation_area=="92040") %>%
  select(percent_of_population_fully_vaccinated)
```


>Q20. Finally make a time course plot of vaccination progress for all areas in the full dataset with a age5_plus_population > 36144.

```{r}
vax.36.all <- filter(vax, age5_plus_population > 36144)


ggplot(vax.36.all) +
  aes(as_of_date, percent_of_population_fully_vaccinated, group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color="blue") +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated",
       title="Vaccination Rate Across California",
       subtitle="Only areas with a population above 36k are shown") +
  geom_hline(yintercept = 0.6629812, linetype="dashed")
```


>Q21. How do you feel about traveling for Thanksgiving and meeting for in-person class next Week?

I think it will be ok, so long as people take the neccessary measures while traveling (masks, vaccination standing, distancing when able, isolation if positive).












