---
title: "class10.Rmd"
author: "Anu Chaparala"
date: "10/28/2021"
output: pdf_document
---

```{r}
url <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/candy-power-ranking/candy-data.csv"
  
candy <- read.csv(url, row.names=1)
head(candy, n=5)
#View(candy)
```

>Q1. How many different candy types are in this dataset?

There are 85 different candy types in this dataset.

```{r}
dim(candy)

#or

nrow(candy)
```

>Q2. How many fruity candy types are in the dataset?

There are 38 fruity candy types in the dataset. 

```{r}
sum(candy$fruity)
```

>Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

Nestle Butterfinger with 70.7 winpercent. 

```{r}
candy["Nestle Butterfinger",]$winpercent
```

>Q4. What is the winpercent value for “Kit Kat”?

76.8 winpercent. 

```{r}
candy["Kit Kat",]$winpercent
```


>Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```

```{r}
#install.packages("skimr")
library(skimr)
skim(candy)
```


>Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

winpercent looks like it's on the 0-100 scale rather than 0-1. 

>Q7. What do you think a zero and one represent for the candy$chocolate column?

True/False values

>Q8. Plot a histogram of winpercent values

```{r}
hist(candy$winpercent, xlab = "Win Percent")
```

>Q9. Is the distribution of winpercent values symmetrical?

No, it is slightly skewed right.

>Q10. Is the center of the distribution above or below 50%?

The center is below 50%.

>Q11. On average is chocolate candy higher or lower ranked than fruit candy?

It appears that on average, chocolatety candy is higher ranked than fruity candy. (61% vs. 44%).

First we need to find all the chocolate candy rows in the 'candy' data. 

```{r}
choc <- as.logical(candy$chocolate)
data_choc <- candy [choc,]$winpercent
mean(data_choc)

fruit <- as.logical(candy$fruity)
data_fruit <- candy [fruit,]$winpercent
mean(data_fruit)

```

>Q12. Is this difference statistically significant?

Yes, this difference is significant with p-value less than 0.05. (2.871e-08)

```{r}
#Run t-test with choc and fruit vectors (data_choc and data_fruit) to check for statistical significance. 
t.test(data_choc, data_fruit)
```

>Q13. What are the five least liked candy types in this set?

Nik L Nip, Boston Baked Beans, Chiclets, Super Bubble, and Jawbusters.

```{r}
head(candy[order(candy$winpercent),], n=5)
```


>Q14. What are the top 5 all time favorite candy types out of this set?

Snickers, Kit Kat, Twix, Reeses Miniatures, and Reeses Peanut Butter Cup.

```{r}
tail(candy[order(candy$winpercent),], n=5)
```

>Q15. Make a first barplot of candy ranking based on winpercent values.


```{r}
library(ggplot2)
ggplot(candy) +
  aes(winpercent, rownames(candy)) +
  geom_col()
```

Improve this plot by reordering by the 'winpercent' value. 
>Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent.

```{r}
ggplot(candy) +
  aes(winpercent, reorder(rownames(candy), winpercent)) +
  geom_col()
```

Let's add some color! Set up a color vector for coloring candy by candy type.

```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"

ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 
```

>Q17. What is the worst ranked chocolate candy?

Sixlet

> Q18. What is the best ranked fruity candy?

Starburst

##PricePercent

```{r}
#install.packages("ggrepel")
library(ggrepel)

# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)

#change fruity color to be visible
my_cols[as.logical(candy$fruity)] = "red"
```


>Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

Reeses Miniatures

>Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

Nik L Nip, Nestle Smarties, Ring pop, Hersheys Krackel, and Hersheys Mild Chocolate. Nik L Nip is most expensive and least popular!

```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```

##Exploring the Correlation Structure

```{r}
#install.packages("corrplot")
library(corrplot)

cij <- cor(candy)
corrplot(cij)
```

>Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

fruity and chocolate

>Q23. Similarly, what two variables are most positively correlated?

chocolate and bar


##PCA

```{r}
pca <- prcomp(candy, scale=TRUE)
summary(pca)
```
```{r}
#Now we can plot our main PCA score plot of PC1 vs PC2.
plot(pca$x[,1:2])
#add color
plot(pca$x[,1:2], col=my_cols, pch=16)
```

```{r}
# Make a new data-frame with our PCA results and candy data
my_data <- cbind(candy, pca$x[,1:3])

p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)

p
```



```{r}

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```

```{r}
#install.packages("plotly")
library(plotly)
ggplotly(p)
```

```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```

>Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

Fruity, hard, and pluribus, which make sense since fruity candies are typically hard and come in a box of multiples (like mikenikes!)



